Leadership across cultures and contexts: an exploratory analysis

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Tweaked mosaic plots

loadd(all_data)
loadd(all_data2)
loadd(all_study_vars)
loadd(coercive_plot)
loadd(df_func)
loadd(df_qual)
loadd(features)
loadd(feature_models)
loadd(feature_models_aic)
loadd(follower_benefits_txt)
loadd(follower_costs_txt)
loadd(follower_benefits_cult)
loadd(follower_costs_cult)
loadd(functions_support_cult)
loadd(functions_support_txt)
loadd(highstatus_plot)
loadd(leader_benefits_txt)
loadd(leader_benefits_cult)
loadd(leader_costs_txt)
loadd(leader_costs_cult)
loadd(m_lpca_func)
loadd(m_lpca_qual)
loadd(m_pvclust_func)
loadd(m_pvclust_qual)
loadd(multi_aic)
loadd(multi_allvars)
# loadd(p_heatmap_groups)
# loadd(p_heatmap_region)
# loadd(p_heatmap_sex)
# loadd(p_heatmap_subsis)
loadd(p_heatmap_feature_subsis)
loadd(p_heatmap_feature_region)
loadd(p_heatmap_feature_sex)
loadd(p_heatmap_feature_groups)
loadd(plot_costs_benefits)
loadd(plot_elastic_status)
loadd(plot_elastic_coercive)
loadd(plot_features_support)
loadd(plot_func_qual)
loadd(plot_pubdate)
loadd(qualities_support_cult)
loadd(qualities_support_txt)

@@

```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group context. Areas of the bars are proportional to the number of text records in that category."}
ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
plot_group_subsis #+ plot_group_sex
```

@@

```{r cluster, fig.cap="Cluster analysis of leader quality variables (A) and leader function variables (B). Distances were $1 - cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

@@

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)). The composite feature measures mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial. See Figure \@ref(fig:featuressupportplot).

@@

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group varition (i.e., the four factors). See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

The leadership "features" we derived by clustering leadership dimensions were represented across most cultures. Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions. Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions. See Figure \@ref(fig:featuressupportplot).

@@

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egaliarian societies.

```{r matingbenefit}
matingbenefit <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable=='Mating benefit'], 3)
```

Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019]. But the computational services model emphasizes mating benefits, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality". See Figure \@ref(fig:varsupportcb).

@@

```{r regionsubsistence, fig.width=11, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}
ggplot(all_data) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
```

>>>>

loadd(all_data)
loadd(all_data2)
loadd(all_study_vars)
loadd(coercive_plot)
loadd(df_func)
loadd(df_qual)
loadd(features)
loadd(feature_models)
loadd(feature_models_aic)
loadd(features_support_cult)
loadd(follower_benefits_txt)
loadd(follower_costs_txt)
loadd(follower_benefits_cult)
loadd(follower_costs_cult)
loadd(functions_support_cult)
loadd(functions_support_txt)
loadd(highstatus_plot)
loadd(leader_benefits_txt)
loadd(leader_benefits_cult)
loadd(leader_costs_txt)
loadd(leader_costs_cult)
loadd(m_lpca_func)
loadd(m_lpca_qual)
loadd(m_pvclust_func)
loadd(m_pvclust_qual)
loadd(multi_aic)
loadd(multi_allvars)
# loadd(p_heatmap_groups)
# loadd(p_heatmap_region)
# loadd(p_heatmap_sex)
# loadd(p_heatmap_subsis)
loadd(p_heatmap_feature_subsis)
loadd(p_heatmap_feature_region)
loadd(p_heatmap_feature_sex)
loadd(p_heatmap_feature_groups)
loadd(plot_costs_benefits)
loadd(plot_elastic_status)
loadd(plot_elastic_coercive)
loadd(plot_features_support)
loadd(plot_func_qual)
loadd(plot_pubdate)
loadd(qualities_support_cult)
loadd(qualities_support_txt)

@@

```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group context. Areas of the bars are proportional to the number of text records in that category."}
ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
plot_group_subsis
```

@@

```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables. Distances were $1-cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

@@

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organize cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligent, and experienced and accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductively, and socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)). The composite feature variables mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, and material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial. See Figure \@ref(fig:featuressupportplot).

@@

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable and intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that does not vary much by the four factors. See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions. Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural pressures, and receiving various benefits in exchange for their functions. The leadership "features" we derived by clustering leadership dimensions were represented across most cultures. See Figure \@ref(fig:featuressupportplot).

@@

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this universal list corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity and status leveling mechanisms.

```{r matingbenefit}
matingbenefit <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable=='Mating benefit'], 3)

reprosuccess <- signif(100*features_support_cult$Estimate[features_support_cult$Variable == 'Social_material_success'], 3)
```

Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019]. The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)). But there was evidence for the "Social material success" feature, which included Polygyny, High quality spouse, and Many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)). Interestingly, the computational model highlights women's family leadership role, and the estimated marginal mean for Social material success was higher for females leaders than male leaders (see Figure \@ref(fig:heatmap_feature_sex)), with the caveat that there were few text records on female leaders in our sample.

@@

```{r regionsubsistence, fig.width=12, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}

lbls <- c(
  'Africa', 
  'Circum-\nMediterranean', 
  'East\nEurasia', 
  'Insular\nPacific',
  'North\nAmerica',
  'South\nAmerica'
  )

ggplot(all_data) + 
  geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) +
  scale_fill_viridis(discrete = T) + 
  scale_x_productlist(labels = lbls) +
  guides(fill = guide_legend(reverse = T)) +
  labs(x = '', y = '') +
  theme_minimal(15)
```
